---
title: "Creating a Kernel"
output: rmarkdown::html_vignette
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Creating a Kernel}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include = FALSE}
library(distr6)
set.seed(42)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Kernel Class

Similarly to implemented probability distributions, implemented kernels all inherit from a single
class 'kernel', which in turn inherits from `Distribution` (see the [uml diagram](https://raw.githubusercontent.com/wiki/alan-turing-institute/distr6/images/uml.png)).  But this is a much simpler class with reduced methods and many of these return the same result and are therefore defined in the `kernel` parent class. For every kernel the following is true

* `kernel$mean()` == 0
* `kernel$median()` == 0
* `kernel$mode()` == 0
* `kernel$type()` == Reals$new()
* `kernel$valueSupport()` == continuous
* `kernel$variateForm()` == univariate

And for all kernels the sampling method uses inverse transform sampling, so this is also defined in
the parent class. Hence the only methods required to add are

* squared2Norm: This is new for kernels (i.e. not given for probability distributions)
* variance

And then the d/p/q functions are given in the constructor just like probability distributions.


## Creating a Kernel

### Kernel Variables

 These are identical to the SDistribution public variables:

* name - Full (unique) name of kernel
* short_name - Short name (unique) id for kernel
* description - Short description, usually just the name
* package - The package in which the d/p/q functions are written

For the Epanechnikov kernel, the above all looks like

```{r eval=FALSE}
Epanechnikov <- R6::R6Class("Epanechnikov", inherit = Kernel, lock_objects = F)
Epanechnikov$set("public","name","Epanechnikov")
Epanechnikov$set("public","short_name","Epan")
Epanechnikov$set("public","description","Epanechnikov Kernel")
```

Note:

1. Again note the use of `lock_objects = F` which ensures correct usage with decorators.
2. The `package` variable is omitted as it defaults to 'distr6'

### Kernel Methods

As stated above, there are fewer methods that need to be implemented in kernels than probability distributions. These include


* variance
* squared2Norm: the squared 2-norm of the kernel's pdf over the full support

Again the d/p/q methods are implemented in the constructor. So for the Epanechnikov kernel,

```{r eval=FALSE}
Epanechnikov$set("public","squared2Norm",function(){
  return(3/5)
})
Epanechnikov$set("public","variance",function(){
  return(1/5)
})
```

Note:

1. These methods will always return constants, they are defined as methods not variables for consistency
2. If no constant numeric is available, omit the method so a numeric one can be added after decoration
3. As well as the methods listen above, the following methods are also included by default and don't need to be defined
    * prec
    * stdev
    * iqr

### The Constructor

The constructor for kernels is much more simple than that of probability distributions and only need include the d/p/q methods and properties. For the Epanechnikov kernel:

```{r eval=FALSE}
Epanechnikov$set("public","initialize",function(decorators = NULL){
  pdf <- function(x1){
    return(0.75 * (1-x1^2))
  }
  cdf <- function(x1){
    return(3/4*x1 - 1/4*x1^3 + 1/2)
  }

  super$initialize(decorators = decorators, pdf = pdf, cdf = cdf,
                   support = Interval$new(-1, 1),  symmetric = TRUE)
  invisible(self)
})
```

Note:

1. We omit the quantile method as no closed form analytic expression was found
2. The support will generally either be `Reals$new()` or `Interval$new(-1,1)`
3. Kernels are all symmetric
4. The constructor always takes one argument only, `decorators` which is passed to the parent-class constructor

## Summary

Kernels are much simpler to extend than SDistributions. Just remember the following

1. The 4 public variables: name, short_name, description, package
2. The 2 public methods: variance, squared2Norm
3. The constructor: Includes d/p/q, the only argument is `decorators`, and the properties are generally identical

## Extension Guidelines

* [SDistribution](https://alan-turing-institute.github.io/distr6/articles/webs/create_sdistribution.html)
* [Wrapper](https://alan-turing-institute.github.io/distr6/articles/webs/create_wrapper.html)
* [Decorator](https://alan-turing-institute.github.io/distr6/articles/webs/create_decorator.html)
