---
title: "Getting and Setting Parameters"
output: rmarkdown::html_vignette
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Getting and Setting Parameters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include = FALSE}
library(distr6)
set.seed(42)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

In our [previous tutorial](https://alan-turing-institute.github.io/distr6/articles/webs/constructing_a_distribution.html) we constructed a Normal distribution with a variety of parameterisations. Now we will look at how to access the parameters of the distribution and how to update them. Parameters are handled by [param6](https://cran.r-project.org/package=param6), full documentation can be found at the package [website](https://xoopr.github.io/param6/).

## Accessing Parameters

First we will construct the Standard Normal distribution, the default parameterisation
```{r}
N <- Normal$new()
```

To view all parameters in the distribution we use the `parameters` method
```{r}
N$parameters()
```

Notice how not only the parameters in the given parameterisation are displayed, but all the ones possible. Individual values can be obtained with `getParameterValue`, which is a wrapper around `get_values`.

```{r}
N$getParameterValue("prec")
```

## Updating Parameters

Any parameter can be updated in distr6 using the `setParameterValue` method and all others are updated accordingly. For example,

```{r}
N$setParameterValue(var = 2)
N$parameters()
```

Not only has the variance been updated, but so too have the precision and standard deviation parameters. But be careful not to set conflicting parameterisations.

The wrong way:

```{r, error=TRUE}
N$setParameterValue(prec = 2)
```

The right way:

```{r}
N$setParameterValue(prec = 2, var = NULL)
```

Internal checks ensure that only valid parameter values are allowed, for example setting the precision to a negative number will throw an error.

```{r,error=TRUE}
N$setParameterValue(prec = -1)
```

Finally, multiple parameters can be updated at the same time

```{r}
N$setParameterValue(mean = 4, prec = 2)
N$parameters()
```

## Summary

In this tutorial we looked at getting and setting parameters for the Normal distribution. In the [next tutorial](https://alan-turing-institute.github.io/distr6/articles/webs/statistical_methods.html) we look at accessing mathematical and statistical methods including the d/p/q/r functions.
