---
title: "Statistical Methods"
output: rmarkdown::html_vignette
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Statistical Methods}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include = FALSE}
library(distr6)
set.seed(42)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

In the [previous tutorial](https://alan-turing-institute.github.io/distr6/articles/webs/parameters.html) we constructed a Normal distribution and accessed and updated its parameters for a variety of parameterisations. In this tutorial we cover how to access mathematical and statistical methods of the Normal distribution including the `dnorm/pnorm/qnorm/rnorm` equivalents in distr6.

## d/p/q/r

The advantage of distr6 over R stats is that once a distribution is constructed, it's very easy to find properties and results from the distribution by changing very little. For simple distributions like the Normal distribution, this may not seem like a big difference, but for more complicated ones with multiple parameters, you'll find yourself saving a lot of time!

Once again we start with constructing the Standard Normal distribution
```{r}
N <- Normal$new()
```

For simplicity, we refer to both the probability density functions of continuous distributions and probability mass functions of discrete distributions, as the "pdf" function. This is in line with R stats using "d" for "density". The other statistical methods from R stats are referred to as "cdf", "quantile" and "rand", the same as in R stats:

```{r}
N$pdf(1:2) # Density evaluated at points '1' and '2'
N$cdf(1:2) # Distribution function evaluated at points '1' and '2'
N$quantile(0.975) # Quantile function evaluated at 0.975
N$rand(5) # 5 samples from the Normal distribution
```

## Other Mathematical and Statistical Methods

We have seen in the [first tutorial](https://alan-turing-institute.github.io/distr6/articles/webs/constructing_a_distribution.html) how the `summary` method can be used to view quick statistics about a probability distribution, i.e.
```{r}
summary(N)
```

But all these statistics can be accessed individually as well. To see the full list of available methods view the 'Statistical Methods' section of the distribution help page, `?Normal`. All probability distributions have the same methods available if possible, i.e. If there is an analytic expression for a statistical result, then we provide it! Below are just a few examples

```{r}
N$mean()
N$variance()
N$skewness()
N$kurtosis(excess = FALSE)
N$cf(2)
N$mgf(2)
```


## Summary

In this tutorial we looked at using the d/p/q/r functions in distr6 and accessing other statistical results. In the [next tutorial](https://alan-turing-institute.github.io/distr6/articles/webs/properties_traits.html) we take a quick look at distribution properties and traits, whilst trying not to get into too big a discussion about object-oriented programming!
