---
title: "Overview to distr6 v0"
output: rmarkdown::html_vignette
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Overview to distr6 v0}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#")
```

## What is distr6?

* distr6 is an object-oriented interface for probability distributions in R
* It adds the following functionality to base R distributions:
  * A unified OOP interface
  * Basic statistical properties (mean, variance,...)
  * More advanced numerical properties (p-norms, survival modelling,...)
  * ParameterSets for flexible parameterisation
  * SymbolicSets for flexible domain representation
  * Wrappers for joint, product, array, truncated,... distributions

## Installing distr6

As distr6 is currently in development, there is no publication to CRAN and therefore it has to be installed via the `remotes` package. Either install the latest stable version from master or the latest dev version:

```{r results = "hide",eval=FALSE}
# remotes::install_github("RaphaelS1/distr6") # Installs stable version from master
remotes::install_github("RaphaelS1/distr6", ref = "dev") # Installs latest dev version
```

Then attach the library as usual:

```{r results = "hide",message=FALSE,warning=FALSE}
library(distr6)
```

## distr6 vs. R Stats

distr6 is the disruptive R packageto takeover base distributions functionality, or at least to add in a few new handy methods! We understand many people come to R for statistics, and probability distributions are a central part of this. With that in mind, let's look at how distr6 has advantages over R stats. First of all, let's say we want to evaluate the mass, distribution and inverse distributions functions for the Binomial distribution.

### R Stats
```{r}
dbinom(1:10, size = 10, prob = 0.5)
pbinom(1:10, size = 10, prob = 0.5)
qbinom(seq.int(0,1,length.out=10),size = 10, prob = 0.5)
```

### distr6
```{r}
B <- Binomial$new(size = 10, prob = 0.5)
# Option 1
B$pdf(1:10)
B$cdf(1:10)
B$quantile(seq.int(0,1,length.out=10))

# Option 2
cdf(B, 1:10)

# Option 3
library(magrittr)
B %>% cdf(1:10)
```

distr6 allows the user to create a distribution object and then call any methods on this object, meaning that the exact parameterisation and parameter values don't have to be remembered. This is especially useful after statistical inference when the parameters can be arbitrarily hard to remember, e.g. `Binomial(size = 100, prob = 0.62938)`.

Moreover once an object is created, the user can call statistical methods to get basic (and more complex) results
```{r}
B$mean()
summary(B)
```

### Listing Distributions

Currently we are planning on implementing around 60 probability distributions for distr6 v1. At least 20 of these are essential. Another advantage of distr6 over R stats is in efficiency to list and choose distributions, this is especially useful for less-technical users who have come to R for distibutions only.

```{r}
listDistributions()
listDistributions(traits = list(valueSupport = "discrete"))
```

## Wrappers and Decorators

Users familiar with OOP may be aware of Design Patterns, code templates used to solve recurring tasks. We make use of 'wrappers' and 'decorators' throughout distr6.
* Decorators are used to add functionality to an object
* Wrappers take this further by adapting the interface and changing the object name.

```{r}
N <- Normal$new()
# N$survivalAntiDeriv() - Produces an error as it currently doesn't exist as a method
decorate(N, ExoticStatistics) # Alternatively: ExoticStatistics$new(N)
N$survivalAntiDeriv(lower = 1, upper = 4)
```

In distr6 we use the principle of minimising numeric results but when used always produce a warning. Finally an example of a wrapper is the truncated distribution:

```{r}
TN = TruncatedDistribution$new(N, lower = 0, upper = 5) # Alternatively: truncate(N, lower = 0, upper = 5)
TN$pdf(6)
TN$pdf(0)
TN$pdf(2)
print(TN)
```

### Listing wrappers and decorators
Again so the user knows what is available, we have functionality to list available decorators and wrappers

```{r}
listWrappers(simplify = TRUE)
listDecorators(simplify = TRUE)
```

## Parameterisation
distr6 assumes that the user knows which parameterisation they want to use for probability distributions but if they do not, an educated guess is made. All common parameterisations are possible, for example for the Normal distribution

```{r}
Normal$new(mean= 0, sd = 1)
Normal$new(mean = 0, var = 1)
Normal$new(mean = 0, prec = 2)
```

In construction, the given parameters are taken to be the parameterisation, so in the above examples the Normal distribution is constructed with the mean and standard deviation, variance or precision respectively. In any case, the other options are automatically updated:
```{r}
N <- Normal$new(mean = 0, sd = 2)
N$parameters()
N$setParameterValue(list(prec = 2)) # Fails as precision isn't a settable parameter
N$setParameterValue(list(sd = 3))
N$parameters()
```

## Future Plans

We plan on implementing the following in distr6 v1:

* 60 probability distributions with the following:
  * Analytic results for basic statistical results including mean, var and more complex modelling results including p-norms and anti-derivatives
* Basic numeric results for the CoreStatistics and ExoticStatistics decorators
* Basic imputation methods for pdf/cdf/quantile/rand
* Wrappers for joint, product, array, truncated, huberized and conditional distributions
