---
title: "Intro to distr6 v1"
output: html_notebook
---

## What is distr6?

* distr6 is intended to (in the best-case scenario) replace probability distributions in R stats
* It adds the following functionality to distributions:
  * A unified OOP interface
  * Basic statistical properties (mean, variance,...)
  * More advanced numerical properties (p-norms, survival modelling,...)
  * ParameterSets for flexible parameterisation
  * SymbolicSets for flexible domain representation
  * Wrappers for joint, product, array, truncated,... distributions

## Installing distr6

As distr6 is currently in development, there is no publication to CRAN and therefore it has to be installed via the `remotes` package. Either install the latest stable version from master or the latest dev version:

```{r eval = FALSE}
remotes::install_github("RaphaelS1/distr6") # Installs stable version from master
remotes::install_github("RaphaelS1/distr6", ref = "dev") # Installs latest dev version
```

Then attach the library as usual:

```{r results = "hide"}
library(distr6)
```

Alternatively if testing code that you are currently working on, use `devtools` once your code has been saved and your working directory is set to `distr6`:

```{r eval=FALSE}
devtools::document() # It's good practice to run this first
devtools::load_all()
```

## distr6 vs. R Stats vs. distr
For any new package to be useful it either has to be a significant improvement on existing ones (if they exist) or introduce substantial and worthwhile contributions. It should also be easy to use or at the very least have sufficient documentation so that a user can learn how to work with the package. Many people come to R for statistics and probability distributions are a central part of this therefore distr6 should also be intuitive. With that in mind, let's look at how distr6 has advantages over R stats. First of all, let's say we want to evaluate the mass, distribution and inverse distributions functions for the Binomial distribution.

### R Stats
```{r}
dbinom(1:10, size = 10, prob = 0.5)
pbinom(1:10, size = 10, prob = 0.5)
qbinom(seq.int(0,1,length.out=10),size = 10, prob = 0.5)
```

### distr6
```{r}
B <- Binomial$new(size = 10, prob = 0.5)
# Option 1
B$pdf(1:10)
B$cdf(1:10)
B$quantile(seq.int(0,1,length.out=10))

# Option 2
#pdf.Distribution(B, 1:10)
cdf(B, 1:10)
quantile(seq.int(0,1,length.out=10))

# Option 3
library(magrittr)
B %>% cdf(1:10)
```

distr6 allows the user to create a distribution object and then call any methods on this object, meaning that the exact parameterisation and parameter values don't have to be remembered. This is especially useful after statistical inference when the parameters can be arbitrarily hard to remember, e.g. `Binomial(size = 100, prob = 0.62938)`.

Moreover once an object is created, the user can call statistical methods to get basic (and more complex) results
```{r}
B$expectation() # via R6
summary(B)
```

### Listing functionality
One of the advantages of defining classs, is how easy it is to list what has been implemented and available to the user. For example,
```{r}
listDistributions()
listWrappers(simplify = TRUE)
listDecorators(simplify = TRUE)
```


# Future plans


